---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(virionData)
library(frictionless)
library(kableExtra)
```


## Check available versions of the dataset

```{r list deposits}
 list_deposit_versions()
  deposit_summary()

```

## Get latest version of the data
```{r get data}
# get data - by default this is the latest version of the data.
get_versioned_data(dir_path = "outputs")
# list files
fs::dir_ls("outputs/15692263")
```


## Read the data

Now that you have the data locally, you can read it! 
The virion files are comma delimited with period decimal markers.

```{r read data, eval=FALSE}

virion <- vroom::vroom(file = "outputs/15692263/virion.csv.gz")

```

## Cite the data

Citing data makes increases reproducibility and incentivizes data sharing. 


```{r citation, results='markup'}

# by default the citation will be generated for the current working version.
# this is set when we run `get_versioned_data`
get_version_citation(style = "modern-language-association")
```



```{r citation-specific, results='markup'}
# we can cite a specific version by providing a zenodo id
get_version_citation(zenodo_id = "15643004",style = "apa")

```

```{r bibtex export}
# sometimes you want a bibtex entry item
export_deposit_bibtex("15643004")

```

### What about the data sources?

The data sources used to create Virion are referenced directly in the deposit metadata.
These can be accessed by retrieving the deposit metadata. 

```{r data sources, results='asis'}

metadata_json_text <- export_deposit_metadata(zenodo_id = "15692263",format = "json",verbose = FALSE)

metadata_list <- jsonlite::fromJSON(metadata_json_text,)

related_identifiers <- metadata_list$metadata$related_identifiers

required_items_filter <- related_identifiers$relation_type$id == "requires"

required_items <- related_identifiers[required_items_filter,]

required_items$resource_type <- required_items$resource_type$id

required_items$relation_type <- "requires"

required_items |>
  kableExtra::kable() |>
  kableExtra::kable_material()

```



## What do the fields mean?

Great question! The datapackage.json file contains field descriptions. 
It is fairly human readable but we can take a closer look using some
built in functions.

For a deeper dive, check out the [frictionless](https://docs.ropensci.org/frictionless/) package.

```{r frictionless}

# this function wraps frictionless functions and extracts 
dict_list <- get_data_dictionary(datapackage_json = "outputs/15692263/datapackage.json")

dict_list$detection_csv
```

Use `purrr::map` to make a full data dictionary for a given deposit.

```{r frictionless tables, results='asis'}

html_tables <- purrr::map(dict_list, function(x){
  html_tables <- kableExtra::kable(x = x) |>
  kableExtra::kable_material()
  
})

list_items <- names(dict_list)

tables <- sprintf("<h3>%s</h3><br>%s",list_items,html_tables) |>
  paste(collapse = "<br>")

cat(tables)

```
