---
title: "Working with Virion 2.0"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{virion_2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(virionData)
```


On 28 April 2026, Virion 2.0 was released with goal of streamlining the processing pipeline. Starting with deposit version `19837096`, the data structure diverges significantly from the [original manuscript](https://journals.asm.org/doi/10.1128/mbio.02985-21) and previous versions of the data. 

The major difference is that taxonomic data, database metadata, and NCBI accession numbers are now stored in separate tables to avoid unnecessary operations on those data and to avoid repeating information in the virion table (data normalization). 

```{r join-tables}

# create a new virion deposit object
virion_deposit <- deposit$new()

# set the working version
virion_deposit$set_working_version(print_citation = FALSE,print_bibtex = FALSE)

# get the data
## check the file names
# virion_deposit$working_files$file_key
# 

tax_table <- virion_deposit$load_remote_csv_file(file_key = "tax_table.csv.gz",col_types = "c")

db_table <- virion_deposit$load_remote_csv_file(file_key = "db_table.csv",col_types = "c")


virion_data <- virion_deposit$load_remote_csv_file(file_key = "virion.csv.gz",col_types = "c")

ncbi_accession <- virion_deposit$load_remote_csv_file(file_key = "ncbi_accession.csv.gz",col_types = "c")

# join host tax data
 virion_data_host <- dplyr::left_join(x = virion_data,
                   y = tax_table |>
                      dplyr::rename_with(~ paste0("Host",
                                                  .x,
                                                  recycle0 = TRUE)
                                         ),
                   by = c("HostTaxHashID"))
 
 # join virus tax data
 virion_data_tax <- dplyr::left_join(x = virion_data_host,
                   y = tax_table |>
                      dplyr::rename_with(~ paste0("Virus",
                                                  .x,
                                                  recycle0 = TRUE)
                                         ),
                   by = c("VirusTaxHashID")) |>
   dplyr::mutate(AssocID = as.character(AssocID))

# join NCBI accession numbers
 
virion_data_ncbi <- dplyr::left_join(
  x = virion_data_tax,
  y = ncbi_accession,
  by = "AssocID"
)

# join database data

virion_data_full <- dplyr::left_join(x = virion_data_ncbi,
                                   y = db_table,
                                   by = "DatabaseVersion" )

# if you need the exact same number of columns you could drop
# host and virus tax hash id columns.

```
