---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(virionData)
library(frictionless)
library(kableExtra)
```


## Check available versions of the dataset

```{r list deposits}
 list_deposit_versions()
  deposit_summary()

```

## Get latest version of the data
```{r get data}
# get data
get_versioned_data(dir_path = "outputs")
# list files
fs::dir_ls("outputs/15692263")
```


## Read the data

Now that you have the data locally, you can read it! 
The virion files are comma delimited with period decimal markers.

```{r read data, eval=FALSE}

virion <- vroom::vroom(file = "outputs/15692263/virion.csv.gz")

```


## What do the fields mean?

Great question! The datapackage.json file contains field descriptions. 
It is fairly human readable but we can take a closer look using the frictionless package.

```{r frictionless}

# this function wraps frictionless functions and extracts 
dict_list <- get_data_dictionary(datapackage_json = "outputs/15692263/datapackage.json")

dict_list$detection_csv
```

Use `purrr::map` to make a full data dictionary for a given deposit.

```{r frictionless tables, results='asis'}

html_tables <- purrr::map(dict_list, function(x){
  html_tables <- kableExtra::kable(x = x) |>
  kableExtra::kable_material()
  
})

list_items <- names(dict_list)

tables <- sprintf("<h3>%s</h3><br>%s",list_items,html_tables) |>
  paste(collapse = "<br>")

cat(tables)

```
